---
title: "PA1_template.Rmd"
output:
  html_document:
    keep_md: yes
    self_contained: no
---

## Loading and preprocessing the data

The activity data is read and loaded in the 

```{r, echo=TRUE}
suppressMessages(library(dplyr))
library(knitr)
setwd("~/Google Drive/Coursera//DataScience/05-repro-research/")
activity <- read.csv("activity.csv", na.strings = "NA")
activity <- transform(activity, date=as.Date(as.character(date)))
```

## What is mean total number of steps taken per day?

```{r, echo=TRUE}
out <- activity  %>% filter(!is.na(steps)) %>% group_by(date) %>% summarize(total=sum(steps))

hist(out$total,xlab="steps", ylab="frequency", main="Steps per day" )

round(mean(out$total),0)
median(out$total)
```

The mean for the steps is 10766 (rounded) and median is 10765. 

## What is the average daily activity pattern?

```{r}

out_t <- activity %>% 
  filter(!is.na(steps)) %>%
  group_by(interval) %>% 
  summarize(total=sum(steps,na.rm=TRUE))

max_steps <- max(out_t$total)
max_interval <- out_t[out_t$total== max_steps,][1]

out_m <- activity %>% 
  filter(!is.na(steps)) %>% 
  group_by(interval) %>% 
  summarize(mean=mean(steps,na.rm=TRUE))

plot(out_m$interval, out_m$mean, type="l", ylab="average steps", xlab="interval", main="Average Steps for 5 min interval")
abline(v=max_interval, col="blue")
legend("topright", c("max steps"), lty=c(1), col=("blue"))
axis(at=835, labels=c("835"), las=2, side=1)

```

The max steps (`r max_steps`) is shown as a verticle blue line plotted in the graph

## Imputing missing values
The Missing values are imputed from average value for the interval
```{r}
missing_values <- dim(activity %>% filter(is.na(steps)))[1]

```
`r missing_values` values are missing.
We draw two plots side by side. The first plot is using imputed missing values (mean) and the second is the plot without missing values.

```{r}

out_j <- activity %>% 
  left_join(out_m, by="interval") %>% 
  mutate (steps1 = ifelse (is.na(steps), round(mean), steps))

out_i <- out_j %>% 
  group_by(date) %>% 
  summarize(total=sum(steps1,na.rm=TRUE))

out_o <- out_j %>%
  group_by(date) %>% 
  summarize(total=sum(steps,na.rm=TRUE))

par(mfrow= c(1,2), mar=c(3,2,2,1))
hist(out_i$total, xlab="total steps", ylab="date", main="Steps per day (imputed steps)")
hist(out_o$total, xlab="total steps", ylab="date", main="Steps per day")

summary(out_i$total)
summary(out_o$total)

```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
library("lattice")
# out_i has imputed data, with steps1 as the new column
out_t <- out_j %>% mutate(wday = ifelse(weekdays(date) %in% c("Sunday", "Saturday"), "Weekend", "Weekday")) %>%
  mutate(wday = as.factor(wday)) %>%
  select(steps1, date, interval, wday)

wd_data <- out_t %>% filter(wday=="Weekday") %>%
  group_by(interval) %>%
  summarize(mean = mean(steps1)) %>%
  mutate(wday="Weekday")

we_data <- out_t %>% filter(wday=="Weekend") %>%
  group_by(interval) %>%
  summarize(mean = mean(steps1)) %>%
  mutate(wday="Weekend")

all <- rbind(wd_data, we_data)

xyplot(mean ~ interval |wday ,data=all, layout=c(1,2) , type="l", xlab="Interval", ylab="Steps", main="Weekend vs Weekday data")
```

